---
title: "Variables and other things"
output:
  pdf_document: default
  html_notebook: default
---

## Loading some data from AirBnB

This is an R notebook.

It is a mixture of text, like this, and code, like the "chunk" below.

It allows you to describe your analysis in the same document as the code to do
your analysis.  You can also embed the results.

The result is that if you give someone a notebook like this, they can repeat
your analysis exactly, or change it, to explore or to extend what you did.


```{r}
airbnb <- read.csv('amsterdam_airbnb.csv')
head(airbnb)
```

## What type of thing do I have?

As we'll see in a bit, `read.csv` is a *function*.  It returns something. We've
given that thing a name, `airbnb`.

We very often want to know what type of thing we have.  We can ask what type of
thing we have with the `class` *function*.

```{r}
class(airbnb)
```

The type of thing that came back from `read.csv` was a *data frame*.

We will come back to this soon.

In the next few sections, we try to analyze what R makes of the code we type.
Gradually you will get used to doing this, but it takes some time.

We have to get used to the way that R thinks.

## A comment

Thinking the way R thinks, part 1.

What does R think, when it sees a line that starts with `#`?

It thinks "This is not for me, I can ignore this".

```{r}
# This doesn't do anything
# Whatever I type, R ignores it.
# I can type code, preceded by #, but it does nothing
# head(airbnb)
```

Notice there is no output.

## Functions

Thinking the way R thinks, part 2.

Here we display the first six rows of the data (frame):

```{r}
# The top six rows of the data frame
head(airbnb)
```

What is R thinking?  First it sees `head`.  It realizes that `head` is the name for something.  What is it the name of?  We can check ourselves:

```{r}
# What type of thing is "head"?
class(head)
```

It's the name of a function.  R goes to get that function, ready to use.

Then R sees the parenthesis `(`.  It realizes that we want to *call* the function that it just found \- the `head` function.

A function is like a recipe with a name.  It has instructions on what to do.
There are ingredients, and there is an output, which is a meal.  In the same
way, the function is a procedure.  It needs ingredients, which are the
*arguments* to the function. It applies the procedure to the ingredients, and
returns something.

R knows that the ingredients (arguments) are inside the brackets.  It looks
there, and finds one argument, `airbnb`. R goes to look what that is, and it finds that it's a *data frame*.

```{r}
# What type of thing is "airbnb"?
class(airbnb)
```

R sends the data frame to the function and applies the procedure. When it has
finished applying the procedure, it brings back the result.  In our case, the
result is the first 6 rows of the data frame.

```{r}
# Apply the "head" function, and show the result
head(airbnb)
```

Whenever you see a name, like `head` followed by parentheses, you know this
code *calls* a *function*, with some *arguments* and returns the result.

## Things that come back from functions, and displaying results

You just saw how we call a function and display a result:

```{r}
head(airbnb)
```

R finds what `head` is (a function), finds what `airbnb` is (a data frame), then passes `airbnb` to the function `head`, and returns the result.

In the chunk above, we just displayed the result.  When R returns a result, but we aren't keeping track of the result, it just show it to us.

"Keeping track" means giving the result a name.  We do that, like this:


```{r}
result <- head(airbnb)
```

Notice, because we gave the result a name, R doesn't display anything.  It has the result now, attached to the *name* `result`.

Can you guess what type of thing `result` is?   Hint: it's the top 6 rows of a
data frame.


```{r}
class(result)
```

If we want to display `result`, we can put `result` on its own on the last line of a chunk.  R gets the thing `result` points to, but then sees that we aren't assigning it to anything, so it display it instead:


```{r}
result
```

## Variables

Above, we have been using *variables*.  A *variable* is a name that point to
some value.  It's a label that we can use to refer to something.  We choose the name to be a good label, so, when we read the label, we have a good idea of thing it refers to.

At the moment `result` is a name that point to the first six rows of the data
frame.

We can make simpler variables, as you have already done in your practical:

```{r}
# An assignment statement
a <- 1
```

Let us go through this slowly.   When R sees this, it thinks:

* Aha, the `<-` means this is an *assignment*.
* I'm expecting a variable name (label) on the left of `<-`.
* I'm expecting *something giving a value* on the right of `<-`
* In all "set the variable `a` to have the value `1`.

We can see the value of `a` by using it on it's own in a line.  This finds the value that `a` refers to, and shows it to us:

```{r}
a
```

In the last list, I use *something giving a value*.  The technical term for this is an *expression*.  An *expression* is something R can run, that results in a value.

For example, consider this:


```{r}
b <- 2 + 3
```

Read this as "set the variable `b` to have \- the result of evaluating `2 + 3`.  What value is this?


```{r}
b
```

What do you think you will get for this?

```{r}
c <- b + a
```

Read this as "set the variable `c` to have \- the result of evaluating `a + b`. `b` evaluates to its value: 5. `a` evaluates to its value: 1.  The whole expression evaluates to 1 + 5, giving the value 6.

```{r}
c
```

## Variables, data frames and columns

You have already seen that we can get the column of a data frame by using the
`$` syntax.

For example, we can display all the values of the `price` column in the `airbnb` data frame, like this:

```{r}
airbnb$price
```

What is R thinking here?   First it looks to see what `airbnb` is, and it finds
that it refers to a data frame.  Next is sees the `$`.  It knows that the `$`
means, get something from the thing on the left.  The thing on the left is the
`airbnb` data frame.  Then it looks to see what is at the right of the `$`.  It
is `price`.  It goes to look for something called `price` in the data frame.
In particular, it goes to look for a `price` column in the data frame.  If it
finds it, R gives us back a new value, which is \- a sequence of all the values
for the `price` column.

What kind of thing is this new value, containing all the columns?

```{r}
my_variable <- airbnb$price
```

```{r}
class(my_variable)
```

What do you think would happen if I looked for a column that was not there?

```{r}
another_variable <- airbnb$not_a_column
```

Let's work through it slowly.  R finds the `airbnb` value; it's a data frame.
It sees the dollar.  It tries to find something called `not_a_column` in the
data frame.  It doesn't find anything matching.  It returns a value to tell us
"I couldn't find anything like that in `airbnb`".

```{r}
another_variable
```

## Some exercises

Now consider these chunks:

```{r}
my_variable <- airbnb$price
```

```{r}
my_variable$property_type
```

What happened there?  Work through the logic above, and explain what happened.

How about:

```{r}
airbnb$my_variable
```

What happened there?  Work through the logic above, and explain what happened.

Back to variables. Consider this:

```{r}
x <- 2
```

```{r}
y <- x * 3
```

```{r}
y
```

So far so good.  Now I do:


```{r}
x <- 4
```

What value does `y` have now?  Look through the logic in the variable section above, and work out what R is doing for each step.  Is `y` the same, or is it different, from the value you saw above?  Explain to your neighbour.
